### 2019-03-20 micro4_2.tex
Micro

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1\ The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U
D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

@@

\end{document}

>>>>

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1 The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U
D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

@@

\section*{4 Productive Tasks}
	Extend the job market screening model that we saw in Lecture 8 to the case in which
	tasks are productive. Assume that a type $\theta$ worker produces $\theta(1 + \mu t)$ units of output
	when her task level is $t$ where $\mu > 0$.
	Identify the SPNE of this game.

	\textbf{Solution}
	SPNE is depicted below.
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig4}
	\end{figure}
	As in usual screening model pooling equilibria are impossible. And separating equilibrium is on the picture. It is indeed equilibria because in this case it cannot be beaten a separating equilibrium as in unproductive task model, because in this equilibrium high type obtains higher level of utility than in any pooling equilibrium. And moreover since indifference curve of high type is tangent to the line $w = \theta_H(1 + \mu t)$ firms cannot simultaneously offer a slightly worse (with negative profit) contract for the low type and slightly better (with positive profit) for low type, because the high type will become strictly worse-off.

	\section*{The Patient}
		Suppose that the hypochondriac patient of question 1 is a professional runner. The
		patient has been training for a very important race for the last two years and he is very
		much looking forward to it. In order to run in this important race, first he should be
		cleared that he is "fit for running" by the Organization of the race. Suppose that when
		the patient meets his doctor, there is also a member of the Organization in the surgery.
		What do you think will happen to the quality of the information transmitted by the
		patient?

		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running. That means parameter $b$ becomes smaller. Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running. And the sportsman understands it. Thus, $b$ becomes smaller and as we can see in Problem 1, expected profits increases, i.e. the quality of information transmitted goes up.

		\section*{6 Absence of Equilibria}
			In the job market screening model that we saw in class, we found that a pooling contract
			was not an equilibrium because a firm could make money by skimming off the good types.
			We also saw that a separating equilibrium is beaten by a pooling contract under some
			conditions. Do you view this result as troubling? How might we be able to alter the
			model to get a more reasonable prediction?

			We should use some refinements i.e. slightly alter the concept of equilibrium. We can use anticipatory (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979 ) equilbrium concepts. The anticipatory equilibrium concept manages to deal with absence of pooling equilibria. Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable. That is, once contracts are withdrawn firms cannot skim off only a good type.

			The concept of reactive equilibrium manages to deal with the second problem. This concept allows firms to add more contracts. Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig. \eqref{fig5}, point 2) makes the initial deviation no longer profitable. Hence the separating equilibrium is not beaten by pooling under this concept.
				\begin{figure}[H]
				\centering
				\includegraphics[width=0.8\textwidth]{plotdraft}
				\caption{}\label{fig5}
			\end{figure}
\end{document}
